Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-01 schemes.tex
SRLNC almost done

\subsection{Random Linear Network Coding}

@@

In \eqref{eq:recoded_packet}, $w_{ij}$ is the coding coefficient used
in the $i$-th recoded packet and assigned to multiply a previously coded
packet $C_j$. These coding coefficients are again uniformly and randomly chosen
from $GF(q)$. Still, each of the $w_{ij}$ are not the appended recoding
coefficients for signalling since the they will only retrieve previously
coded packets. The analytical expression for the appended coefficients
after recoding is obtained by calculating recoded packets as function of the
original packets. Therefore, we define the local coding matrix $\textbf{W}$
in the same way as it was made for $\textbf{V}$. Thus, the local coding matrix
can be written as:
%
\begin{align} \label{eq:local_coding_matrix}
\textbf{W} =
\left[
\begin{array}{c}
        w_1    \\ \hline
        \vdots \\ \hline
        w_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        w_{11} & \ldots & w_{1g} \\
        \vdots  & \ddots & \vdots  \\
        w_{g1} & \ldots & w_{gg} \\
\end{array}
\right]
\end{align}
%
With the definitions from \eqref{eq:coding_matrix} and
\eqref{eq:local_coding_matrix}, the recoded packets
$\textbf{R} = \left[R_1 \ldots R_g \right]^T$ are written as
$\textbf{R} = (\textbf{W} \textbf{V}) \textbf{P}$. Here, we recognize
the relationship between original and recoded packets. The resulting
coding matrix is simply the multiplication of matrices $\textbf{W}$ and
$\textbf{V}$. Denoting ${(\textbf{W} \textbf{V})}_{ij}$ as the element
in the $i$-th row and $j$-th column of $(\textbf{W} \textbf{V})$, this
term is the resulting coding coefficient used to create $R_i$ after
encoding the original packet $P_j$ recoding it locally in an intermediate
node. Finally, the appended coefficients for $R_i$ are
${(\textbf{W} \textbf{V})}_{ik}$ with $k \in [1,g]$. By doing some
algebra, each ${(\textbf{W} \textbf{V})}_{ij}$ term can be verified to be
computed as:
%
\begin{align}
\label{eq:appended_coded_coefficients}
{(\textbf{W} \textbf{V})}_{ij} = \sum_{k=1}^{g} w_{ik} v_{kj}\ ,\ \forall i,j \in [1,g] \times [1,g]
\end{align}
%
This procedure is carried by all the recoders in a given network, therefore
keeping the structure of the recoded packets for any decoder to compute
the Gaussian elimination. Similarly as with encoding, any decoder that
collects a set of $g$ \ac{l.i.} recoded packets with their respective
${(\textbf{W} \textbf{V})}_{ij}$, will be able to decode the data as mentioned before in Section \ref{sssec:decoding}.

@@

In \ac{SRLNC}, instead of considering all the packets to create a coded
packet like \ac{RLNC} does, here the encoder sets more coding coefficients to
zero when creating a new coded packet. In practice this can be implemented in
a couple of ways.

%One way to accomplish this is to add a Bernoulli random
%variable, $\text{B}(d)$, as in \ref{eq:sparse_coded_packet}:
The simplest way to accomplish this is to add a Bernoulli random
variable, $\text{B}(d)$, as in \ref{eq:sparse_coded_packet}:

\begin{align} \label{eq:sparse_coded_packet}
    %C_i  &= \bigoplus_{j=1}^{g} \text{B}(d)_{j} \, v_{ij} \otimes P_j ,\ \forall i \in [1,g] ,\
    C_i  &= \bigoplus_{j=1}^{g} \text{B}(d) \, v_{ij} \otimes P_j ,\ \forall i \in [1,g] ,\
    v_{ij} > 0,\
    d \in
    \begin{cases}
        {(0,0.5]} & \text{for } q = 2 \\
        {(0,1]} & \text{for } q > 2
    \end{cases}
\end{align}

We call $d$ the coding density.
Reducing $d$ enables the encoder to decrease the average number of packets that
are mixed into each coded packet. Thereby reducing the complexity of the
encoder that needs to mix less packets, but also on the decoder that in
contrast needs to eliminate less nonzero coding coefficients. The catch of this
approach is that coded packets produced by the encoder becomes more \ac{l.d.}
on each other as the coding complexity is reduced. This incurs in additional
channel overhead.

Generating a coded packet does not imply that it contains any data at all. I.e.
$ \left[ B(d)_{1}, \ldots, B(d)_{g} \right] = \mathbf{0}$.
In that case, one would have to trash the coded packet and generate a new one.

Most coded packets that are produced by the encoder are typically very sparse.
We will therefore rely on a slightly different implementation in KODO
\ref{eq:sparse_coded_packet2} that proved faster for low densities. It is best
expressed as:

\begin{align} \label{eq:sparse_coded_packet2}
    C_i  &= \bigoplus_{j=1}^{\floor{\text{Binom}(g,d)}} v_{im} \otimes P_{m} ,\
    %L \subseteq \{1,\dots,g\}: |L|=k ,\
    \forall i \in [1,g] ,\
    v_{ij} > 0
\end{align}
where
$m = \text{choice}(\{1,\dots,g\}, \text{size}=k, \text{replace}=\text{False})$.
\fxnote{Can this be represented smarter?}

%To solve this, one could trash the coded packet and generate a new coded
%packet. This is not the most efficient because KODO needs to count the
%the number of nonzero coding coefficients in each coded packets to eliminate
%all zero packets.
%
%This is the most correct approach, but the most efficient approach although it is mathematically .

>>>>

\subsection{Random Linear Network Coding}
\label{ssec:RLNC}

@@

In \eqref{eq:recoded_packet}, $w_{ij}$ is the coding coefficient used
in the $i$-th recoded packet and assigned to multiply a previously coded
packet $C_j$. These coding coefficients are again uniformly and randomly chosen
from $GF(q)$. Still, each of the $w_{ij}$ are not the appended recoding
coefficients for signalling since the they will only retrieve previously
coded packets. The analytical expression for the appended coefficients
after recoding is obtained by calculating recoded packets as function of the
original packets. Therefore, we define the local coding matrix $\textbf{W}$
in the same way as it was made for $\textbf{V}$. Thus, the local coding matrix
can be written as:
%
\begin{align} \label{eq:local_coding_matrix}
\textbf{W} =
\left[
\begin{array}{c}
        w_1    \\ \hline
        \vdots \\ \hline
        w_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        w_{11} & \ldots & w_{1g} \\
        \vdots  & \ddots & \vdots  \\
        w_{g1} & \ldots & w_{gg} \\
\end{array}
\right]
\end{align}
%
With the definitions from \eqref{eq:coding_matrix} and
\eqref{eq:local_coding_matrix}, the recoded packets
$\textbf{R} = \left[R_1 \ldots R_g \right]^T$ are written as
$\textbf{R} = (\textbf{W} \textbf{V}) \textbf{P}$. Here, we recognize
the relationship between original and recoded packets. The resulting
coding matrix is simply the multiplication of matrices $\textbf{W}$ and
$\textbf{V}$. Denoting ${(\textbf{W} \textbf{V})}_{ij}$ as the element
in the $i$-th row and $j$-th column of $(\textbf{W} \textbf{V})$, this
term is the resulting coding coefficient used to create $R_i$ after
encoding the original packet $P_j$ recoding it locally in an intermediate
node. Finally, the appended coefficients for $R_i$ are
${(\textbf{W} \textbf{V})}_{ik}$ with $k \in [1,g]$. By doing some
algebra, each ${(\textbf{W} \textbf{V})}_{ij}$ term can be verified to be
computed as:
%
\begin{align}
\label{eq:appended_coded_coefficients}
{(\textbf{W} \textbf{V})}_{ij} = \sum_{k=1}^{g} w_{ik} v_{kj}\ ,\ \forall i,j \in [1,g] \times [1,g]
\end{align}
%
This procedure is carried by all the recoders in a given network, therefore
allowing any decoder to compute the original data after Gaussian elimination,
regardless of the amount of times where recoding was performed. Similarly as
with the encoding operation, any decoder that collects a set of $g$ \ac{l.i.} recoded
packets with their respective coefficients, will be able to decode the data
as mentioned before in Section \ref{sssec:decoding}.

@@

In \ac{SRLNC}, instead of considering all the packets to create a coded
packet as in \ac{RLNC}, for this coding scheme an encoder sets more
coding coefficients to zero when generating a coded packet. A way to
accomplish this, is to set a fixed ratio of non-zero coefficients
in the encoding vector of size $g$. We refer to this fixed ratio as the
average coding density $d$. Thus, for any coded packet $C_i$ with coding
coefficients $v_{ij},\ j \in [1,g]$, its coding density is defined as follows:
%
\begin{align}
\label{eq:coding_density}
d = \frac{\sum_{j=1}^{g} f(v_{ij})}{g},\ \
    f(v_{ij}) =
    \begin{cases}
        0 & \text{for } v_{ij} = 0 \\
        1 & \text{for } v_{ij} \neq 0
    \end{cases}
\end{align}
%
From the density definition in \eqref{eq:coding_density}, it can be observed that
$0 \leq d \leq 1$. As $g$ increased, we obtain more granularity in the density.
Notice that the special case $d = 0$ has no practical purpose since it
implies just zero padded data. So, a practical range for the density excludes
the zero case, $0 < d \leq 1$. Moreover, it can be also observed that the
case of $d = 1$ reduces to the \ac{RLNC} coding scheme described in Section
\ref{ssec:RLNC}.

To achieve a desired average coding density in the coefficients, for each
of them, we utilize a set of Bernoulli random variables all with parameter $d$, i.e.
$\mathbb{B}_j \sim Bernoulli(d)\ ,\ \forall j \in [1,g]$. In this way, we
can represent a coded packet in \ac{SRLNC} as:
%One way to accomplish this is to add a Bernoulli random
%variable, $\text{B}(d)$, as in \ref{eq:sparse_coded_packet}:
\begin{align} \label{eq:sparse_coded_packet}
    %C_i  &= \bigoplus_{j=1}^{g} \text{B}(d)_{j} \, v_{ij} \otimes P_j ,\ \forall i \in [1,g] ,\
    C_i  &= \bigoplus_{j=1}^{g} \mathbb{B}_j v_{ij} \otimes P_j ,\ \forall i \in [1,g] ,\ \
    d \in
    \begin{cases}
        {(0,0.5]} & \text{for } q = 2 \\
        {(0,1]} & \text{for } q > 2
    \end{cases}
\end{align}
%
In \eqref{eq:sparse_coded_packet}, we have specified a dependency on the field size
for practical density values. In case of employing $GF(2)$, the maximum plausible
density is restricted up to 0.5 since higher values incur in more frequent \ac{l.d.}
coded packets \textbf{Include reference here}. Reducing $d$ enables the encoder to
decrease the average number of packets manipulated to make a coded packet. This
reduces the complexity of the encoder since it needs to mix less packets. Moreover,
it also simplifies the decoder processing given that less nonzero coding coefficients
required to be operated on during the Gaussian elimination stage.

The drawback of this scheme is that coded packets from the encoder become more \ac{l.d.}
on each other as the density is reduced. This incurs in transmission overhead
since more coded packets are required to be sent to overcome the reception of redundant
packets. Also, particularly with this method, generating a coded packet may imply
that it could not contain any data at all. For example, we might find the case
where $\left[\mathbb{B}_1, \ldots, \mathbb{B}_g \right] = \mathbf{0}$. In that case,
the encoder discards the coded packet and tries to generate a new one, though this is
unlikely to occur in general.

Under the current method for setting the density, most coded packets that
are produced by the encoder are typically very sparse.
The reason being that here we control the \textit{average} amount
of nonzero coefficients, but not the exact value for each coded packet. Therefore,
we may have some cases where we have more or less packets combined. A second
implementation in Kodo considers doing something more clever.

As is shown in \eqref{eq:sparse_coded_packet2}, that implementation
proved faster for low densities \textbf{reference?}. It is best expressed as:
%
\begin{align} \label{eq:sparse_coded_packet2}
    C_i  &= \bigoplus_{j=1}^{\floor{\text{Binom}(g,d)}} v_{im} \otimes P_{m} ,\
    %L \subseteq \{1,\dots,g\}: |L|=k ,\
    \forall i \in [1,g] ,\
    v_{ij} > 0
\end{align}
where
$m = \text{choice}(\{1,\dots,g\}, \text{size}=k, \text{replace}=\text{False})$.
\fxnote{Can this be represented smarter?}
